---
title: "Figure 2 a&c"
author: "Meng Li"
date: "2023-05-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#setwd("D:/OneDrive - The Pennsylvania State University/Box Sync/Penn_State/Research_data/Seminal_roots_2021/Peng_data/Peng_pop_structure")

#devtools::install_github("bcm-uga/TESS3_encho_sen")
library(tess3r)
library(dplyr)
library(tidyr)
library(tidyverse)
library(data.table)
library(ggplot2)
library(ggpubr)
library(cowplot)
library(tibble)
library(ggh4x)
library(raster)
library(rworldmap)
library(gstat)
library(rnaturalearth)
library(sp)

```

For paralleling computing

```{Parellel_Computing}
#detectCores() #16
library(parallel)
library(snow)
library(doSNOW)
cl <- makeCluster(16, type = "SOCK")
registerDoSNOW(cl)

```

# 1. read orignial and predicted phenotypic data and coordinates
```{r phenotype}
#environment data
pheno <- read.csv("data/FOAM_EnvVars_SRN.csv",header = TRUE, row.names = 'X',na.strings="")
#original phenotypic data
SRN_orig <- read.csv('data/PengData_EnvVars_CleanedImputed_SRN.csv', row.names = 'X')
#RF predicted phenotypic data
SRN <- read.csv("output/SRN_Predictions861Accessions_82322_Final.csv", header = TRUE, row.names = 'X',na.strings="")

pheno_SRN <- pheno %>% 
    dplyr::select(Accession, Elevation) %>%
    inner_join(SRN, by = "Accession") %>%
    dplyr::select(SaID, Longitude, Latitude, SRN_pred,Elevation) %>%
    drop_na(SaID)

pheno_SRN_names <- pheno_SRN$SaID

```

# 2. Visualize SRN map + marginal plot (Figure 2a)

```{r map}
#remove duplicate locations before creating a spatial object (SRN is the same at the same location)
SRN_pheno.1 = pheno_SRN[which(!duplicated(pheno_SRN[2:3])), ]

#convert SRN dataframe into a spatial points object
coordinates(SRN_pheno.1) = ~Longitude+Latitude
plot(SRN_pheno.1)

#Create a grid with a slightly larger extent
#use the locator to click 4 points beyond the extent of the plot
#and use those to set your x and y extents
#locator(4)
lon.range <- as.numeric(c(-120,-80))
lat.range <- as.numeric(c(10,34))
# now expand SRN range to a grid with spacing that to use in the interpolation
#here we will use 0.1 grid cells:
grd <- expand.grid(x=seq(from=lon.range[1], to=lon.range[2], by=0.1), y=seq(from=lat.range[1], to=lat.range[2], by=0.1))
## convert grid to SpatialPixel class
coordinates(grd) <- ~ x+y
gridded(grd) <- TRUE
plot(grd, cex=1.5)
points(SRN_pheno.1, pch=1, col='red', cex=1)

#Now we have set up our points and a grid to interpolate onto, we are ready carry out some interpolation. 
#Kriging requires the construction of a semivariogram model to describe the spatial autocorrelation pattern for your particular variable.
#The values in the cloud can be binned into lags with and plotted with a very similar function
semivariog<-variogram(SRN_pred~1, locations=SRN_pheno.1, data=SRN_pheno.1)
plot(semivariog)
semivariog
#From the empirical semivariogram plot and the information contained in the semivariog gstat object, we can estimate the sill, range and nugget to use in our model semivariogram.
#Range: the point on the distance axis where the semivariogram starts to level off: 9
#Sill: the point on the y axis where the semivariogram starts to level off: 0.5
#Nugget: 0.19
#Using this information we’ll generate a model semivariogram using the vgm() function in gstat.
#first check the range of model shapes available in vgm
vgm()
#try different models and use the one fits the best
#model.variog<-vgm(psill=0.72, model='Wav', nugget=0.19, range=9)
model.variog<-vgm(psill=0.5,model='Wav',nugget=0.19, range=9)
fit.variog<-fit.variogram(semivariog, model.variog)
plot(semivariog, fit.variog)
#using ordinary kriging
krig<-krige(formula=SRN_pred ~ 1, locations=SRN_pheno.1, newdata=grd, model=model.variog)
plot(krig)
save(krig, file = 'output/Final_SRN_interpolation_krige_object.Rimage')

#load('Saved_Rimage_object/Final_SRN_interpolation_krige_object.Rimage')
#create a mask layer 
mex_sp <- ne_countries(country = c("Mexico",'Guatemala','Belize','El Salvador','Honduras'), ##make an sp object of countries
                       returnclass = c("sp")) 

P_r <- raster::raster(krig)
P_r.m <- raster::mask(P_r, mex_sp)
plot(P_r.m)

#dataframe for ggplot
P_r.m_df <- raster::as.data.frame(P_r.m, xy=T)
P_r.m_df <- P_r.m_df %>% dplyr::filter(!is.na(var1.pred))
glimpse(P_r.m_df)

# filter points to the selected area
rownames(SRN_orig) <- SRN_orig$Accession

countriesSP <- getMap(resolution='low')
pointsSP = SpatialPoints(SRN_orig[,c("Longitude","Latitude")], proj4string=CRS(proj4string(countriesSP)))  

# use 'over' to get indices of the Polygons object containing each point 
indices = over(pointsSP, mex_sp)

SRN_orig$name <- indices$name

SRN_orig <- drop_na(SRN_orig, name)

# ggplot map
map.polygon <- getMap(resolution = "low")

# color usa, cub, and nic as white
ct <-  map_data('world') %>%
    filter(region %in% c('USA','Cuba','Nicaragua')) %>%
    mutate(color = 'white')


library(ggExtra)
library(gridExtra)
library(cowplot)

map_box <- ggplot() +
  geom_tile(data = P_r.m_df, aes(x = x, y = y, fill = var1.pred))+
  geom_path(data = map.polygon, aes(x = long, y = lat, group = group)) +
  coord_sf(xlim = c(-120, -82), ylim = c(12, 34), expand = FALSE) +
  geom_point(data = SRN_orig, aes(x = Longitude, y = Latitude),fill = 'white',color = 'black',size = 0.5,shape= 21,stroke = 0.4) +
  #scale_color_gradient(low="white", high="deepskyblue4", name = 'Observed SRN',breaks=c(2,4,6,8)) +
  #geom_polygon(data = ct, aes(x = long, y = lat, group = group), fill = 'white',color = 'black' ) +
  scale_fill_scico(palette='bamako', direction = -1,name = 'Predicted SRN',breaks=c(3,4,5),
                   guide = guide_colorbar(direction = "horizontal",barwidth = unit(2,'cm'), barheight = unit(0.2,'cm'))) +
  force_panelsizes(rows = unit(4, "cm"),cols = unit(6, "cm")) +
    theme_bw()+
    theme(
      #axis.line = element_line(colour = "black", size=0.5), #axis.line or not
      axis.line = element_blank(), # y axis is blank
      #axis.line.x=element_line(colour="black",size=0.5),
      #axis.line.y=element_line(colour="black",size=0.5),
      #axis.ticks = element_blank(),
      panel.border = element_rect(colour="black",size=0.5),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      #axis.text.y = element_text(size = 12,color = 'black',vjust = 0.2,face = 'bold',family = 'sans'), #family = 'sans'
      #axis.text.x = element_text(size=12, color = "black",vjust = 0.2,face = 'bold', family = 'sans'), #for separate plots
      #axis.text = element_blank(),
      #axis.title.y = element_text(size=14, color = 'black',vjust = 1,face = 'bold', family = 'sans'), #for separate plots
      axis.title.y = element_blank(),
      #strip.text = element_text(size=14),
      #axis.title.x = element_text(size =14, vjust = 0.5,color = 'black',face = 'bold', family = 'sans'),
      axis.title.x = element_blank(),
      #plot.tag = element_text(size = 18,color = 'black',face = 'bold', family = 'sans'),
      #plot.tag.position = c(0.1, 1),
      #legend.title = element_text(size=12, color = 'black',vjust = 1,face = 'bold', family = 'sans'),
      legend.title = element_blank(),
      legend.position= c(0.2,0.1),
      #legend.direction="horizontal",
      legend.text=element_text(size=10, color = 'black',vjust = 2,family = 'sans'),
      #legend.key.size = unit(0.5,'cm'),
      #legend.key.width = unit(0.05, 'cm'),
      #legend.key = element_blank(),
      #legend.key = element_rect(colour = "transparent", fill = "transparent")
      legend.background=element_rect(fill = alpha("white", 0)),
      #legend.box.margin=margin(-20,-20,-20,-20),
      #plot.title=element_text(size=18,hjust = 0.5), #face="italic"
      #plot.margin = unit(c(1,1,0,0), "lines") # for separate figures
      #plot.margin = unit(c(0.5,0,0,0), 'lines')
)

map_box


density_ybox <- axis_canvas(map_box, axis = "y", coord_flip = TRUE) + 
      geom_smooth(data = pheno_SRN, aes(x = Latitude, y = SRN_pred,color = 'forestgreen', fill = '#C0CA33'), stat = 'smooth',linewidth = 1) +
      geom_smooth(data = SRN_orig, aes(x= Latitude, y = SRN_avg,color = 'orange', fill = 'wheat'), stat = 'smooth',linewidth = 1 ) +
      scale_colour_manual(name= '',values=c('forestgreen', 'orange'), labels = c("Predicted", "Observed")) +
      scale_fill_manual(name= '',values=c('#C0CA33', 'wheat'), labels = c("Predicted", "Observed")) +
      scale_x_continuous() +
      force_panelsizes(cols = unit(2, "cm"),rows = unit(4, "cm")) + #
      coord_flip() +
      #coord_cartesian(xlim=c(1,5)) + 
    theme_bw()+
      theme(
      #axis.line = element_line(colour = "black", size=0.5), #axis.line or not
      #axis.line = element_blank(), # y axis is blank
      axis.line.y=element_blank(),
      axis.line.x=element_line(colour="black",size=0.5),
      axis.ticks.y = element_blank(),
      panel.border = element_rect(colour="black",size=0.5),
      #panel.border = element_blank(),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      axis.text.x = element_text(size = 12,color = 'black',vjust = 0.2,face = 'bold',family = 'sans'), #family = 'sans'
      #axis.text.x = element_text(size=12, color = "black",vjust = 0.2,face = 'bold', family = 'sans'), #for separate plots
      axis.text.y = element_blank(),
      #axis.title.y = element_text(size=14, color = 'black',vjust = 1,face = 'bold', family = 'sans'), #for separate plots
      axis.title.y = element_blank(),
      #strip.text = element_text(size=14),
      #axis.title.x = element_text(size =14, vjust = 0.5,color = 'black',face = 'bold', family = 'sans'),
      axis.title.x = element_blank(),
      #plot.tag = element_text(size = 18,color = 'black',face = 'bold', family = 'sans'),
      #plot.tag.position = c(0.1, 1),
      #legend.title = element_text(size=12, color = 'black',vjust = 1,face = 'bold', family = 'sans'),
      #legend.title = element_blank(),
      legend.position= 'right',
      legend.direction="horizontal",
      #legend.direction="horizontal",
      #legend.text=element_text(size=10, color = 'black',vjust = 2,family = 'sans'),
      #legend.key.size = unit(0.5,'cm'),
      #legend.key.width = unit(0.05, 'cm'),
      #legend.key = element_blank(),
      #legend.key = element_rect(colour = "transparent", fill = "transparent")
      #legend.background=element_rect(fill = alpha("white", 0)),
      #legend.box.margin=margin(-20,-20,-20,-20),
      #plot.title=element_text(size=18,hjust = 0.5), #face="italic"
      #plot.margin = unit(c(1,1,0,0), "lines") # for separate figures
      #plot.margin = unit(c(0.5,0,0,0), 'lines')
)

p2 <- insert_yaxis_grob(map_box, density_ybox, grid::unit(1, "in"), position = "right")

```

# 3. Calculate pop structure + map magic parents (Figure 2c)

3.1 load highly filtered genotype data
I will try with the highly filtered genotype data; this was used to compute kinship matrix in the GWAS analysis

```{r genotype}
# test with highly filtered genotype
load('genotype/HighFilteredReduced.Rimage')
#genotype data was obtained from:  https://github.com/danjgates/AdaptationScripts2

# change genotype to 0,1,2
X<-genoMDS*2
X<-as.matrix(X)

```

3.2 get needed SRN and coordinates data

```{r need}
# get the genotype names 
inters <-intersect(pheno_SRN_names,rownames(X))

# get the genome data
X_geno<-X[inters,]

# get coordinates
pheno_SRN.1<- pheno_SRN %>% column_to_rownames(var = 'SaID')

coordinates <- pheno_SRN.1[inters,1:2]
coordinates<-as.matrix(coordinates)

# get SRN and coordinates
SRN_pheno <- pheno_SRN.1[inters, 1:4]

```

3.3 estimate ancestry coefficients using tess3; pick the number of groups

```{r coef}
tess3.obj <- tess3(X = X_geno, coord = coordinates, K = 1:9, 
                   method = "projected.ls", ploidy = 2, openMP.core.num = 16) 

save(tess3.obj, file='output/Final_tess3_obj_origin.Rimage')
#load('output/Final_tess3_obj_origin.Rimage')

plot(tess3.obj, pch = 19, col = "blue",
     xlab = "Number of ancestral populations",
     ylab = "Cross-validation score")
# k =5 or 6
# retrieve tess3 Q matrix for K = 5,6 clusters 
q.matrix_5 <- qmatrix(tess3.obj, K = 5)
#q.matrix_6 <- qmatrix(tess3.obj, K = 6)

my.colors <- c("tomato", "orange", "lightblue", "wheat","olivedrab",'purple','darkblue')
my.palette <- CreatePalette(my.colors, 9)

map.polygon <- getMap(resolution = "low")
                      
# color usa, cub, and nic as white
ct <-  map_data('world') %>%
    filter(region %in% c('USA','Cuba','Nicaragua')) %>%
    mutate(color = 'white')
  
pl_5 <- ggtess3Q(q.matrix_5, coordinates, map.polygon = map.polygon, col.palette = my.palette)
pop_5<-pl_5 +
  #ggplot() +
  geom_path(data = map.polygon, aes(x = long, y = lat, group = group)) +
  #xlim(-16, 42) + 
  #ylim(35, 65) + 
  #coord_equal() + 
  coord_sf(xlim = c(-120, -82), ylim = c(12, 34), expand = FALSE) +
  geom_polygon(data = ct, aes(x = long, y = lat, group = group, fill = color),color = 'black' ) +
  scale_fill_identity() +
  force_panelsizes(rows = unit(4, "cm"),cols = unit(6, "cm")) + # set the panel size
  #scale_color_gradient(low = 'white',high = 'black', #space = 'Lab', #'#57C7EB'
  #                         name='Predict SRN') +
  #xlab("Longitute") +
  #ylab("Latitude") + 
  theme_bw() +
    theme(legend.key.size = unit(0.08,'cm'),
            legend.key.width = unit(0.08, 'cm'),
            #legend.box.margin=margin(-10,0,0,-10),
            legend.title = element_text(size=4, color = 'black', family = 'sans'),
            #legend.title = element_blank(),
            legend.text=element_text(size=3,color = 'black',vjust = 0.5, family = 'sans'),
            #legend.direction = "vertical", 
            #legend.box = "horizontal",
            #legend.text = element_blank(),
            axis.ticks = element_blank(),
            axis.text.y = element_blank(),
            axis.text.x = element_blank(),
            #axis.title.y = element_text(size =4, vjust = 0.5,color = 'black',face = 'bold', family = 'sans'),
            axis.title = element_blank(),
          #axis.title.x = element_text(size =4, vjust = 0.5,color = 'black',face = 'bold', family = 'sans'),
            plot.margin = unit(c(1,0,1,0), "lines"),
            panel.grid.major = element_blank(),
            #panel.grid.major = element_line(color = gray(.5), 
            #                                linetype = "dashed", 
            #                                size = 0.1), 
            panel.background = element_rect(fill = 'white'))
pop_5
```

3.4 plot magic founder

```{r magic}

magic_founder <- read.csv('data/MAGIC_Founders_color.csv', header = TRUE)
magic_founder$Race <- factor(magic_founder$Race)


library(scico)
library(ggh4x) # for changing panel size
# Change plot size to 4 inch x 3inch
#options(repr.plot.width=10, repr.plot.height=4)
SRN_map<- ggplot() +
  #geom_tile(data = P_r.m_df, aes(x = x, y = y, fill = var1.pred))+
  geom_path(data = map.polygon, aes(x = long, y = lat, group = group)) +
  coord_sf(xlim = c(-120, -82), ylim = c(12, 34), expand = FALSE) +
  #geom_point(data = SRN_orig.1, aes(x = Longitude, y = Latitude),fill = 'white',color = 'black',size = 0.5,shape= 21,stroke = 0.4) +
  #scale_color_gradient(low="white", high="deepskyblue4", name = 'Observed SRN',breaks=c(2,4,6,8)) +
  geom_point(data=magic_founder,aes(x = Longitude, y = Latitude,group = Race, color = Race), size =1.5, shape = 19) +
  scale_color_manual(values = c("#5EB7BE", "#337C61", "#E1C34A", "#D8904F", "#725E5E", "#F0B3A2", "#88CF56", "#A44638")) +
  #geom_polygon(data = ct, aes(x = long, y = lat, group = group), fill = 'white',color = 'black' ) +
  force_panelsizes(rows = unit(4, "cm"),cols = unit(6, "cm")) + # set the panel size
  theme_bw() +
    theme(legend.key.size = unit(0.3,'cm'),
            legend.key.width = unit(0.15, 'cm'),
            legend.box='horizontal',
            legend.margin=margin(0,0,0,0),
            legend.box.margin=margin(-10,-10,-10,-10),
            legend.title = element_text(size=3, color = 'black', family = 'sans'),
            #legend.title = element_blank(),
            legend.text=element_text(size=3,color = 'black',vjust = 0.5),
            #legend.position = c(0.25,0.15),
            legend.position = 'bottom',
            #legend.direction = "vertical", 
            #legend.box = "horizontal",
            #legend.text = element_blank(),
            axis.ticks = element_blank(),
            axis.text.y = element_blank(),
            axis.text.x = element_blank(),
            #axis.title.y = element_text(size =4, vjust = 0.5,color = 'black',face = 'bold', family = 'sans'),
            axis.title = element_blank(),
          #axis.title.x = element_text(size =4, vjust = 0.5,color = 'black',face = 'bold', family = 'sans'),
            plot.margin = unit(c(0,0,0,0), "cm"),
            panel.grid.major = element_blank(),
            #panel.grid.major = element_line(color = gray(.5), 
            #                                linetype = "dashed", 
            #                                size = 0.1), 
            panel.background = element_rect(fill = 'white'))

SRN_map

```